---
title: "COPEP - COVID19 cluster RCT simulations"
author: "Mark Jones"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    css: "rep.css"
    fig_caption: yes
classoption: landscape  
bibliography: [book.bib]
biblio-style: apalike  
---


<!--

  

output: distill::distill_article
-->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(
  suppressWarnings(library(tidyverse)))
suppressPackageStartupMessages(
  suppressWarnings(library(gridExtra)))
suppressPackageStartupMessages(
  suppressWarnings(library(latex2exp)))
suppressPackageStartupMessages(
  suppressWarnings(library(kableExtra)))
suppressPackageStartupMessages(
  suppressWarnings(library(gridExtra)))
suppressPackageStartupMessages(
  suppressWarnings(library(flextable)))

source("report.R")
source("decision.R")

options(knitr.kable.NA = '') 

dirc = "outrar"

```



```{r, echo = F}

include_rar4 <- T
include_extra <- F

ld1 <- get_results(dirc = "outrar")

ld1$dtrial <- ld1$dtrial %>%
  dplyr::rename(eff = sup,
                 ineff = inf_or_equ)

darm1 <- ld1$darm %>% 
  dplyr::select(-arms_in_post) %>%
  dplyr::group_by(scen, arm, enabled_for_anly, p_tru) %>%
  dplyr::summarise(p_est = mean(p_est, na.rm = T),
                   p_sd = sqrt(mean(p_var, na.rm = T)),
                   active = mean(active, na.rm = T),
                   sup_at = mean(sup_at, na.rm = T),
                   inf_at = mean(inf_at, na.rm = T), 
                   equ_at = mean(equ_at, na.rm = T), 
                   is_best = mean(is_best, na.rm = T),
                   is_eff = mean(is_sup, na.rm = T),
                   is_inf = mean(is_inf, na.rm = T),
                   is_equ = mean(is_equ, na.rm = T),
                   nk = mean(nk, na.rm = T),
                   nki = mean(nki, na.rm = T)) %>%
  dplyr::ungroup() %>%
  dplyr::rename(eff_at = sup_at,
                ineff_at = inf_at,
                is_ineff = is_inf,
                enabled_from = enabled_for_anly) %>%
  dplyr::mutate(enabled_from = enabled_from - 1)

# flextable doesn't like nan
darm1$eff_at[is.nan(darm1$eff_at)] <- NA
darm1$ineff_at[is.nan(darm1$ineff_at)] <- NA
darm1$equ_at[is.nan(darm1$equ_at)] <- NA

if(include_rar4){
  ld2 <- get_results(dirc = "outrar4")
  
  ld2$dtrial <- ld2$dtrial %>%
  dplyr::rename(eff = sup,
                 ineff = inf_or_equ)

  darm2 <- ld2$darm %>% 
    dplyr::select(-arms_in_post) %>%
    dplyr::group_by(scen, arm, enabled_for_anly, p_tru) %>%
    dplyr::summarise(p_est = mean(p_est, na.rm = T),
                   p_sd = sqrt(mean(p_var, na.rm = T)),
                   active = mean(active, na.rm = T),
                   sup_at = mean(sup_at, na.rm = T),
                   inf_at = mean(inf_at, na.rm = T), 
                   equ_at = mean(equ_at, na.rm = T), 
                   is_best = mean(is_best, na.rm = T),
                   is_eff = mean(is_sup, na.rm = T),
                   is_inf = mean(is_inf, na.rm = T),
                   is_equ = mean(is_equ, na.rm = T),
                   nk = mean(nk, na.rm = T),
                   nki = mean(nki, na.rm = T)) %>%
  dplyr::ungroup() %>%
  dplyr::rename(eff_at = sup_at,
                ineff_at = inf_at,
                is_ineff = is_inf,
                enabled_from = enabled_for_anly) %>%
  dplyr::mutate(enabled_from = enabled_from - 1)
  
  # flextable doesn't like nan
  
  darm2$eff_at[is.nan(darm2$eff_at)] <- NA
  darm2$ineff_at[is.nan(darm2$ineff_at)] <- NA
  darm2$equ_at[is.nan(darm2$equ_at)] <- NA
}


# ld3 <- get_results(dirc = "outfix")
# 
# darm3 <- ld3$darm %>% 
#   dplyr::select(-arms_in_post) %>%
#   dplyr::group_by(scen, arm, enabled_for_anly, p_tru) %>%
#   dplyr::summarise(p_est = mean(p_est, na.rm = T),
#                    p_sd = sqrt(mean(p_var, na.rm = T)),
#                    active = mean(active, na.rm = T),
#                    sup_at = mean(sup_at, na.rm = T),
#                    inf_at = mean(inf_at, na.rm = T), 
#                    equ_at = mean(equ_at, na.rm = T), 
#                    p_beat_soc = mean(p_beat_soc, na.rm = T),
#                    p_equ_soc = mean(p_equiv_soc, na.rm = T),
#                    is_best = mean(is_best, na.rm = T),
#                    is_eff = mean(is_sup, na.rm = T),
#                    is_inf = mean(is_inf, na.rm = T),
#                    is_equ = mean(is_equ, na.rm = T),
#                    nk = mean(nk, na.rm = T)) %>%
#   dplyr::ungroup()

# flextable doesn't like nan
# darm3$sup_at[is.nan(darm3$sup_at)] <- NA
# darm3$inf_at[is.nan(darm3$inf_at)] <- NA
# darm3$equ_at[is.nan(darm3$equ_at)] <- NA

if(include_extra){
  ld4 <- get_results(dirc = "outfix4")

  darm4 <- ld4$darm %>% 
    dplyr::select(-arms_in_post) %>%
    dplyr::group_by(scen, arm, enabled_for_anly, p_tru) %>%
    dplyr::summarise(p_est = mean(p_est, na.rm = T),
                   p_sd = sqrt(mean(p_var, na.rm = T)),
                   active = mean(active, na.rm = T),
                   sup_at = mean(sup_at, na.rm = T),
                   inf_at = mean(inf_at, na.rm = T), 
                   equ_at = mean(equ_at, na.rm = T), 
                   p_beat_soc = mean(p_beat_soc, na.rm = T),
                   p_equ_soc = mean(p_equiv_soc, na.rm = T),
                   is_best = mean(is_best, na.rm = T),
                   is_eff = mean(is_sup, na.rm = T),
                   is_inf = mean(is_inf, na.rm = T),
                   is_equ = mean(is_equ, na.rm = T),
                   nk = mean(nk, na.rm = T)) %>%
  dplyr::ungroup()
  
  # flextable doesn't like nan
  darm4$sup_at[is.nan(darm4$sup_at)] <- NA
  darm4$inf_at[is.nan(darm4$inf_at)] <- NA
  darm4$equ_at[is.nan(darm4$equ_at)] <- NA
}



```


## Trial context

Investigating the prophylactic benefit of chemoprevention for post-exposure to an index-case patient who has confirmed COVID-19 disease. 
Evaluating preventative/prophylactic measures in an attempt to reduce the spread of COVID-19 and/or reduce symptom severity in participants who become infected from contact with an index-case patient.

Arms:

1. SOC
2. Hydroxychloroquine
3. Interferon

With more arms possibly introduced over time.

## Overview

The design is an adaptive cluster RCT. 
Close contacts, which at the moment is just eligible household members, that are associated with an index case are identified and define to be a cluster (the index case is not considered part of the cluster).
Each cluster is randomised to a treatment arm or SOC.
At the time of writing, the Australian pandemic status has low numbers of daily cases and there are some restrictions that limit social mixing. 
In such a setting, we anticipate cluster sizes to small; smaller than ideal.
Nevertheless, our simulations have assumed small cluster sizes and so our evaluation of operating characteristics is conservative. 
Should societal restrictions be eased, which they probably will be, we anticipate that what might reasonably considered to be a cluster of close-contacts could increase in size.
Of course, this is assuming we are able to adequately trace all the contacts in a cluster.

Adaptive elements to the design include response adaptive allocation (RAR), the ability to add new arms, sequential analysis and pre-specified decisions for how effective, ineffective and equivalent treatments impact the status of the trial.

The primary endpoint is confirmed SARS-CoV-2 detection from any PCR specimen at day 10 after randomisation (i.e. a binary variable), but for the sake of brevity, we will just use *infected* as representative of an event.
In our evaluation of infections, we consider all members of the cluster regardless of what their (likely unknown) infection status was at the time of entry into the trial.

When conducting an interim analysis, we note that some participants in the data available at the time of the interim will be enrolled, but will not have reached their endpoint. 
Currently, we do not impute these data but rather analyse only the observed data points.
Finally, treatments are evaluated independently, i.e. interaction effects are not currently considered.

We assume that three arms are active at the start of the trial (and for these simulations).
However, the trial can be extended up to any number of arms, subject to the limits of operational practicality and funding.
RAR is enabled following a burn in period of equal allocation to each arm.

The model used for all analyses is a logistic regression with varying intercept for each cluster and parameterised with terms for the intercept and treatment effects associated with each arm.
The model gives estimates for the log-odds of infection and log-odds-ratios for treatment effects.

All comparisons are made solely relative to the SOC until a treatment arm is deemed effective. 
We use the term *effective* to mean there is evidence to strongly suggest that an arm under consideration is better than a comparator arm, typically the SOC.
So *effective* basically has the same semantics of how *superior* would be used in a frequentist trial. 

We make the distinction because the concept of *superiority* has an implicit symmetry with *inferiority*.
That is, if arm 1 is *superior* to arm 2, then by definition, arm 2 is *inferior* to arm 1.
By defining and referring to the concept of *effectiveness*, we de-emphasise binary thinking and talk about and make decisions based more broadly on the strength of evidence available and our level of acceptable risk for making a false decision. 
For example, we might determine a treatment to be *effective* beyond our levels of doubt if there is greater than 0.95 probability that it is less than a comparator. 
Moreover, we might determine a treatment to have a low chance of *effeciveness* and not warrant further exploration, if there is less than 0.2 probability that it is less than a comparator.
Purely for convenience we will refer to such an arm as being *ineffective* and refrain from using the term *inferior* unless the above mentioned symmetry holds.

Another term that we use is *equivalence* which is defined as in terms of the probability mass of the distribution of the difference between the infection rate for any two arm being within a given pre-specified margin.
In lay terms, *equivalence* literally aligns with the common conception of two treatments having the same observed infection response.

The use of the term *futile* in relation to a treatment arm is reserved purely for those cases where the posterior predictive distribution is examined.
As we do not use predictive probabilities in the copep trial, we do not use the term *futile* as a designation of treatment status.

If a treatment arm is deemed effective, we test whether the remaining treatment arms are inferior to the effective arm and deactivate those that prove to be so.
The rationale for doing this is that we consider it unethical to allocate patients to treatments that are inferior to those that have evidence of effectiveness.
Otherwise, we continue with the trial, collecting evidence to make subsequent decisions on the remaining arms.
As the SOC is implicitly inferior to an effective arm, we stop allocation to SOC, but continue to make comparisons relative to the SOC.
We suspend the trial only if there are no further comparisons to be made, i.e. we have reached a conclusion on all arms.
Sometimes a trial will not stop at any of the interim analyses in which case the trial would run to its maximum sample size and then proceed to a final analysis.

As noted, at each interim we fit a GLMM (a logistic regression model with random intercept to account for clustering) and use the joint posterior to determine several probability metrics for making decisions. 
Letting $\theta_k$ be the estimated log-odds of positive PCR results at day 10 and letting $k=1$ represent the SOC arm, we compute:

+ the probability of an arm being better than control; $Pr(\theta_k < \theta_{1})$ $\forall \ k > 1$
+ the probability of an arm being best; $Pr(\theta_k > \theta_j)$ $\forall \ k \ne j$
+ the probability that each treatment arm is equivalent to SOC; $Pr(|\theta_1 - \theta_j|<\delta)$ $j > 1$ 
+ the probability that each remaining active arms is inferior to arms determined to be effective; $Pr(\theta_a > \theta_{eff})$ $\forall \ a$ active arms

Then we compare these values to predetermined decision thresholds and update the trial and arm status accordingly.  

## Simulations

The three-arm scenarios we considere:

+ Null case (suffix "-0") `p_tru` = (0.15, 0.15, 0.15)
+ One effective (suffix "-1") `p_tru` = (0.15, 0.1, 0.15)
+ One very effective (suffix "-2") `p_tru` = (0.15, 0.085, 0.15)
+ Two effective with gradient (suffix "-3") `p_tru` = (0.15, 0.1, 0.125)
+ Two effective equivalent (suffix "-4") `p_tru` = (0.15, 0.1, 0.1)

The four-arm scenarios (with the fourth arm entering at the first interim) we consider:

+ Null case (suffix "-0") `p_tru` = (0.15, 0.15, 0.15, 0.15)
+ One effective (suffix "-1") `p_tru` = (0.15, 0.1, 0.15, 0.15)
+ One very effective (suffix "-2") `p_tru` = (0.15, 0.085, 0.15, 0.15)
+ Three effective with gradient (suffix "-3") `p_tru` = (0.15, 0.1, 0.125, 0.11)
+ Three effective with one very effective (suffix "-4") `p_tru` = (0.15, 0.1, 0.1, 0.085)

## Results

### Commentary and points of note



### Trial-level summaries

Trial level characteristics table headings:

+ `scen` - simulation scenario (see above)
+ `nmax` - maximum number of clusters
+ `nint` - maximum number of interim analyses
+ `intat` - interval between interim analyses
+ `sig` - standard deviation used to generate the random effect
+ `thresh_sup` - probability threshold to declare effective arm
+ `thresh_fut` - probability threshold to declare futile arm
+ `thresh_equ` - probability threshold to declare equivalent arm within +/- 0.5 log-odds
+ `stop` - proportion of trials stopped for effectiveness, futility or equivalence (all arms futile/equiv)
+ `eff` - proportion of trials where effective arm identified
+ `ineff` - proportion of trials deemed ineffective (translates to either ineffective or equivalent)
+ `no_decision` - proportion of trials where neither of the above conclusions were reached
+ `pb` - proportion of trials where best arm identified based on probability of being best
+ `fa` - proportion of trials running all the way to the final analysis
+ `na` - expected number of interim analyses
+ `nk` - expected total number of clusters
+ `nki` - expected total number of participants
+ `durn` - expected duration in days
+ `nsim` - number of simulations of trial specification
+ `muc` - expected cluster size


```{r, echo = F}
gen_trial_table <- function(dtmp){
  
  shade_cols <- names(dtmp)[-1]

  shade_rows <- seq(1, nrow(dtmp), by = 2)
  
  myft <- flextable(dtmp) 
  myft <- theme_vanilla(myft)
  
  myft <- colformat_num(x = myft, 
                        j = c("thresh_sup", "thresh_fut", "thresh_equ", "sig", 
                              "stop", "eff", "ineff", "no_decision", "pb", "fa"),
                        big.mark=",", 
                        digits = 3, 
                        na_str = "")
  
  myft <- colformat_num(x = myft, 
                        j = c("muc"),
                        big.mark=",", 
                        digits = 2, 
                        na_str = "")
  
  myft <- colformat_num(x = myft, 
                        j = c("na", "nk", "nki", "durn"),
                        big.mark=",", 
                        digits = 1, 
                        na_str = "")
  myft <- bg(x = myft, 
             i = shade_rows,
             j = shade_cols,
             bg = "#e0e0e0")
  myft <- autofit(myft)
  
  myft
  
}

```


#### Response adaptive randomisation (3 Arm)

<div style="margin-bottom:50px;">
```{r, layout="l-screen-inset", echo = F, include = T, eval = T}
dtmp <- ld1$dtrial %>% dplyr::select(-fname) 

myft <- gen_trial_table(dtmp)
myft <- set_caption(myft, "Summary operating characteristics for 3 arm RAR design")

myft
```
</div>




#### Response adaptive randomisation (4 arm)

<div style="margin-bottom:50px;">
```{r, layout="l-body-outset", echo = F, include = include_rar4, eval = include_rar4}
dtmp <- ld2$dtrial %>% dplyr::select(-fname) 

myft <- gen_trial_table(dtmp)
myft <- set_caption(myft, "Summary operating characteristics for 4 arm RAR design")

myft
```
</div>

#### Fixed size 3 arm trial

<div style="margin-bottom:50px;">
```{r, echo = F, include = F, eval = F}
dtmp <- ld3$dtrial %>% dplyr::select(-fname) 

myft <- gen_trial_table(dtmp)
myft <- set_caption(myft, "Summary operating characteristics for 3 arm fixed size design")

myft
```
</div>

#### Fixed size 4 arm trial

<div style="margin-bottom:50px;">
```{r, echo = F, include = include_extra, eval = include_extra}
dtmp <- ld4$dtrial %>% dplyr::select(-fname) 
myft <- gen_trial_table(dtmp)
myft <- set_caption(myft, "Summary operating characteristics for 4 arm fixed size design")

myft
```
</div>

### Arm-level summaries

Arm level characteristics table headings:

+ `scen` - simulation scenario (see above)
+ `arm` - treatment arm (1 = SOC)
+ `enabled_from` - time at which arm activated (in interim analyses)
+ `p_tru` - true underlying probability of response
+ `p_est` - expected probability of response (based on posterior)
+ `p_sd` - expected sd of probability of response
+ `active` - proportion of trials where arm was active
+ `eff_at` - expected number of clusters on arm when identified as effective
+ `ineff_at` - expected number of clusters on arm when identified as ineffective
+ `equ_at` - expected number of clusters on arm when identified as equivalent
+ `is_eff` - proportion of trials where arm identified as effective
+ `is_ineff` - proportion of trials where arm identified as futile
+ `is_equ` - proportion of trials where arm identified as equivalent to soc 
+ `nk` - expected number of clusters on arm
+ `nki` - expected number of participants on arm

If a trial is activated at time 0 then it is activated from the start of the trial.
A trial activated at time 1 will be allocated participants after the first interim and then will enter the analyses and decision framework at the second analysis.

```{r, echo = F}
gen_arm_table <- function(dtmp){
  
  shade_cols <- names(dtmp)[-1]
  shade_rows <- seq(1, nrow(dtmp), by = 2)
  myft <- flextable(dtmp) 
  myft <- theme_vanilla(myft)
  myft <- colformat_num(x = myft, 
                        j = c("p_tru", "p_est", "p_sd"),
                        big.mark=",", 
                        digits = 3, 
                        na_str = "")
  myft <- colformat_num(x = myft, 
                        j = c("active", "is_best", "is_eff", "is_ineff", "is_equ"),
                        big.mark=",", 
                        digits = 2, 
                        na_str = "")
  myft <- colformat_num(x = myft, 
                        j = c("eff_at", "ineff_at", "equ_at", "nk", "nki"),
                        big.mark=",", 
                        digits = 1, 
                        na_str = "")
  myft <- bg(x = myft, 
             i = shade_rows,
             j = shade_cols,
             bg = "#e0e0e0")
  myft <- autofit(myft)
  myft
}
```

#### Response adaptive randomisation (3 Arm)

<div style="margin-bottom:50px;">
```{r, echo = F, include = T, eval = T}
myft <- gen_arm_table(darm1)
myft <- set_caption(myft, "Arm-level operating characteristics for 3 arm RAR design")
myft
```
</div>

#### Response adaptive randomisation (4 arm)

<div style="margin-bottom:50px;">
```{r, echo = F, include = include_rar4, eval = include_rar4}
myft <- gen_arm_table(darm2)
myft <- set_caption(myft, "Arm-level operating characteristics for 4 arm RAR design")
myft
```
</div>

#### Fixed size 3 arm trial

<div style="margin-bottom:50px;">
```{r, echo = F, include = F, eval = F}
myft <- gen_arm_table(darm3)
myft <- set_caption(myft, "Arm-level operating characteristics for 3 arm fixed design")
myft
```
</div>

#### Fixed size 4 arm trial

<div style="margin-bottom:50px;">
```{r, echo = F, include = include_extra, eval = include_extra}
myft <- gen_arm_table(darm4)
myft <- set_caption(myft, "Arm-level operating characteristics for 4 arm fixed design")
myft
```
</div>






# Appendix

Formula for ICC

$\text{ICC} = \frac{\sigma_{C}^2}{\sigma_{C}^2 + (\pi^2)/ 3}$
